Testing and Isolation Efficacy:\\Insights from a Simple Epidemic Model

### 2021-07-14 SIR_manuscript.tex

The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g. \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]). This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction. A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE. If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again). If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity. 
We can show that this phenomenon \emph{only} occurs under targeted testing ($w_I > w_S$ ), but we have not yet found a simple explanation of why it cannot occur without it. This phenomenon is also unlikely to occur in the real world. In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).

>>>>

The other counterintuitive result from our analysis is that, for sufficiently high testing intensity $\rho$, further increasing testing intensity can actually \emph{increase} $\Rnum$ (e.g., \fref{pan2}(b), upper right panel [$\theta\_c=1$, $\theta\_w=0$]). This phenomenon can occur because we are considering the DFE in the presence of testing; thus there is an equilibrium distribution of susceptibles between the $S\_n$ (untested) and $S\_u$ (waiting) compartments even as the disease approaches extinction. A higher rate of testing leads to a greater proportion of individuals waiting for negative tests at the DFE. If infected, individuals in this group will take longer to be tested again and to subsequently isolate (because they must wait for their negative tests to be returned before being tested again). If isolation in this group ($\theta\_w$) is low, this effect can under some (relatively rare) circumstances (high $\theta\_c$, low $\omega$, high $\rho$) allow $\Rnum$ to increase with testing intensity. 
We can show that this phenomenon occurs \emph{only} under targeted testing ($w_I > w_S$), but we have not yet found a simple explanation of why it cannot occur without it. This phenomenon is also unlikely to occur in the real world. In particular, it depends on levels of testing that are unrealistically high (at least in large, general-population settings).
